---
title: "40-modeling"
output: html_notebook
---

# Modeling

We'll be using H2O to provide high-speed model solutions and interactive model exploration. 

## Prepare data

We'll add features to the dataset based on the ideas from the group. 

```{r}
attrition_fe_csv <- "attrition_fe.csv"

attrition_fe <- attrition %>%
  education_rel_department() %>% 
  monthy_income_rel_department() %>%
  years_since_last_promotion_rel_department() %>%
  percent_salary_hike_rel_department() %>%
  write_csv(attrition_fe_csv)
```

*Cut and paste the section below to the Future Work slide*

## Future Work

Attrition based on the performance rating.

factors of interest include:
size of team, team dynamics
gender, (manager and the match of the gender) female/male manager
compentation of the ratio and the market
promotion is correlative to the expection (generational) 
flexibility of remote/wfh
work status:contract/salary/hourly based
commuting time

30% diversity


## Run models

We'll call H2O, split the data, and run models. We'll simplify by applying just Random Forest models, and a GLM to examine contribution of specific levels of factors of variables. 


```{r}
data <- h2o.importFile(attrition_fe_csv)
data$attrition <- as.factor(data$attrition)  #encode the binary repsonse as a factor
h2o.levels(data$attrition)
y <- "attrition"
x <- setdiff(names(attrition_fe), c(y))

```

## Split data

Partition the data into training, validation and test sets

```{r}
splits <- h2o.splitFrame(data = data, 
                         ratios = c(0.7),  #partition data into 70%, 15%, 15% chunks
                         seed = 1)         #setting a seed will guarantee reproducibility
train <- splits[[1]]
valid <- splits[[2]]
```

## Run models

```{r}
rf_fit2 <- h2o.randomForest(x = x,
                            y = y,
                            training_frame = train,
                            model_id = "rf_fit2",
                            validation_frame = valid,  #only used if stopping_rounds > 0
                            ntrees = 100,
                            seed = 1)
```


## Calculate value

Given the estimates from the attendees in the earlier session, determine what the value of solution would have been had we had it in place over the time represented in the data. 


```{r}
attrition_salary <- attrition_fe %>%
  select(attrition, hourly_rate) %>% # Need only these fields for cost calcs
  mutate(
    annual_salary = 2080 * hourly_rate,  # 2080 is the number of hours worked a year when full time
  ) 

attrition_predict <- as.data.frame(h2o.predict(rf_fit2, data)) # Predict entire dataset (note that we are not using out-of-bag samples)


#Test 
attrition_cost <-
  cbind(attrition_salary, attrition_predict) %>%                                             # Predict uses position only
  mutate(replace_cost = if_else(attrition == "Yes", replace_multiplier * annual_salary, 0),  # Calc replacement cost
  expected_value_model = if_else(                                                            # Find expected value (see below)
    attrition == "Yes" & predict == "Yes",
    replace_cost * (1 - success_rate) * Yes, # DOUBLE CHECK THIS
    if_else(attrition == "Yes" & predict == "No", replace_cost, 0)
  )) %>%
  summarise(no_model = sum(replace_cost), model = sum(expected_value_model)) 



savings <- (attrition_cost$no_model - attrition_cost$model)/1000000

glue('\n\n\nModel would have saved: ${savings}M')

```

Baseline model (no engineered features): $17.23495M

*NOTE: Need number of employees retained*

*Note on calculating expected value*
The prediction for each employee comes with a probability which can be read as the probability an employee just like this one will leave. 
